---
title: "03_NumberConcentrationRescaling"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{R Initialize, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(readxl)
library(openxlsx)
library(stringr)
library(knitr)
library(scales)

source("paths.R")
source("R/CorrectionFactor.R")
```

Create folders for figures if they do not yet exist

```{R}
concentrations_dir <- paste0(output_dir, "Figures/Rescaled")

if (!dir.exists(concentrations_dir)) {
  dir.create(concentrations_dir, recursive = TRUE)
}

alphas_dir <- paste0(output_dir, "Figures/Alphas")

if (!dir.exists(alphas_dir)) {
  dir.create(alphas_dir, recursive = TRUE)
}
```

# Import data

## Load concentrations

The below data is read in from raw data provided by respective data owners. This is done using ParticleDataPrep.Rmd.

```{r}
DataVersionConc <- "2026-07-07"

# if data is not available, then unzip this file
# unzip(zipfile = paste0(project_dir, output_dir,"Data.zip"),
      # exdir = output_dir)

concentrations <- readRDS(paste0(project_dir,
                                 output_dir,
                                 "Data/",
                                 DataVersionConc,"Particle_concentrations_total.rds"))

# Exclude samples from Mintenig (2020)  which were indicated only for MP > 300 um.
concentrations <- concentrations |>
  filter(!endsWith(Location_name,"*"))
```

## Load alphas

The powerlawfit.csv files are generated using the parallel.Rmd script. These are the results from a HPC run using said script.

TODO: please provide versions for the powerlawFit and powerlawFitLDIR which correspond to versions of the Parallel script. Also, the Output directory is confusing here.

```{r}
DataVersionPowerlawFTIR <- "2026-06-30"

alphasFITR <- read.csv(paste0(project_dir,
                              output_dir,
                              "/Data/powerlawFitFTIR_All",
                              DataVersionPowerlawFTIR,".csv"))

DataVersionPowerlawLDIR <- "2026-04-09"

alphasLDIR <- read.csv(paste0(project_dir,
                            output_dir,
                            "Data/powerlawFitLDIR_All",
                            DataVersionPowerlawLDIR,".csv"))
```

Join together and denote which is FTIR and which LDIR (could probably do this in the exporting in parallel.Rmd)

```{r}
alphas <- alphasFITR|>
  mutate(Analysis_method = "ATR-FTIR and Micro-FTIR")|>
  bind_rows(alphasLDIR|>
              mutate(Analysis_method = "LDIR"))

rm(alphasFITR,
   alphasLDIR)  
```

# Rescale concentrations

Below we rescale the reported concentrations per river.

## Per measurement

### Join concentrations with detection limits

Join concentrations and detection limits etc. in with the alphas and filter out aggregated datasets.

> Note by joining this way we lose all samples from Leslie because no Alpha was derived for this (it has no corresponding river)

```{r}
ConcPSD <- concentrations|>
  filter(!is.na(River))|>
  left_join(alphas |> # leaves out the ones where id == All
              select(id,
                     param,
                     alpha,
                     alpha_lower,
                     alpha_upper,
                     Analysis_method),
            by = c("River" = "id",
                   "Analysis_method" = "Analysis_method"),
            relationship = "many-to-many")|>
  filter(!Analysis_method=="Light microscopy")

rm(concentrations)
```

### Perform rescaling

We rescale each measurement using the alpha for the applicable water body. We distinguish between LDIR and FTIR measurements and use alphas derived from fits using data corresponding to the analysis method. (LDIR data uses an LDIR alpha)

The uncertainty in the fit of the alpha and uncertainty in the Lmin and Lmax are included as follows:

- A triangular distribution is used based on the fitted alpha and its CI
- A triangle distribution is used for Lmin and Lmax with a 20% uncertainty, with a minimum of 1 micrometer and maximum of 5000 micrometer.

```{r Uncertain data}
# Function for creating the uncertain data
fAlignUncertain <- function(
    ConcAlphaCombined, # data frame/tibble with specific concentration and alpha data
    nRuns = 100, # number of runs
    SizeLimitLowerUmin = 0.1, #fraction af increase of the size limit
    SizeLimitLowerUmax = 0.1, #fraction af increase of the size limit
    SizeLimitUpperUmin = 0.1, #fraction af increase of the size limit
    SizeLimitUpperUmax = 0.1, #fraction af increase of the size limit
    SizeParameter = "longest" # can be longest or shortest for fit of alpha
    ){ #

    ConcAlphaCombined |>
    filter(param == SizeParameter) |> 
    mutate(LowerSamplingSizeLimit_um_Tlowl = 
             ifelse(((1-SizeLimitLowerUmin)*Lower_sampling_size_limit_um) > 1,
                      ((1-SizeLimitLowerUmin)*Lower_sampling_size_limit_um), 1),
           LowerSamplingSizeLimit_um_Tupl = 
             (1+SizeLimitLowerUmax)*Lower_sampling_size_limit_um,
           UpperSamplingSizeLimit_um_Tlowl = 
             (1-SizeLimitUpperUmin)*Upper_sampling_size_limit_um,
           UpperSamplingSizeLimit_um_Tupl =
             ifelse(((1+SizeLimitUpperUmax)*Upper_sampling_size_limit_um) < 5000,
                    ((1+SizeLimitUpperUmax)*Upper_sampling_size_limit_um), 5000)
    ) |> 
    rowwise()|>
    mutate(
  alpha_lower = ifelse(alpha < alpha_lower, alpha, alpha_lower),
  alpha_upper = ifelse(alpha > alpha_upper, alpha, alpha_upper),
  check_alpha = (alpha_lower <= alpha & alpha <= alpha_upper)
) |>
    mutate(
      UncertainVariables = list(tibble(
        triangle_alpha = triangle::rtriangle(
          n = nRuns,
          a = alpha_lower,
          b = alpha_upper,
          c = alpha
        ),
        triangle_sampling_lower =
          triangle::rtriangle(
            n= nRuns,
            a= LowerSamplingSizeLimit_um_Tlowl,
            b= LowerSamplingSizeLimit_um_Tupl,
            c= Lower_sampling_size_limit_um
          ),
        triangle_sampling_upper = triangle::rtriangle(
          n= nRuns,
          a= UpperSamplingSizeLimit_um_Tlowl,
          b= UpperSamplingSizeLimit_um_Tupl,
          c= Upper_sampling_size_limit_um
        ),
        UncId = c(1:nRuns)
      ))
    )|> unnest(UncertainVariables)
}

ConcAlphaUncertain <- 
  fAlignUncertain(
    ConcAlphaCombined = ConcPSD, 
    nRuns = 1000,
    SizeParameter = "longest")
```

### Align using triangle distribution in alpha

Because the correction factor (CF) is not calculated linearly we use Monte Carlo simulations to approximate the error.

We pull a triangular distribution of alphas from the measured alpha and the 5th and 95h CI. Then we calculate a corrected value for each iteration.

> We cannot use the bootstrapped alpha values here to derive the uncertainty as they are only used to derive the uncertainty and test the accuracy of the fit. It is also the reason you do not use alphaMean

We also already prepare here the triangle distribution of the upper and lower sampling limits using a 20% spread of the reported value and a triangle distribution.

Calculate the correction factor per simulation:

```{R Align using triangle distribution in alpha}
ConcAlphaAlignedUncertain <-
  ConcAlphaUncertain |> 
  mutate(CF_option2 = # not used in the report
           F.Correctionfactor(Upper_sampling_size_limit_um, # fixed Upper
                              Lower_sampling_size_limit_um, # fixed Lower
                              5000,
                              1,
                              triangle_alpha))|>
  mutate(concentrationOption2 = CF_option2*
           concentration) |> 
  mutate(CF_option3 = 
           F.Correctionfactor(triangle_sampling_upper,
                              triangle_sampling_lower,
                              5000,
                              1,
                              triangle_alpha))|>
  mutate(concentrationOption3 = CF_option3*
           concentration)

ConcAlphaAlignedUncertain_L <-
  ConcAlphaAlignedUncertain |> 
  rename(reported = concentration) |> 
  pivot_longer(
    cols = c(reported,concentrationOption2,concentrationOption3),
    names_to = "OptionType",
    values_to = "concentration"
  )

rm(ConcAlphaUncertain)

ConcAlphaAlignedUncertain_L |> distinct(Data_source)
```

Optional: save the rescaled concentrations

```{R Save rescaled concentrations, eval=FALSE, include=FALSE}
# saveRDS(ConcAlphaAlignedUncertain_L, paste0(project_dir, output_dir, "Data/Rescaled_concentrations_", DataVersionConc, ".Rds"))
```

# Prepare for plotting

## Define plot themes

### Violin plots

Define variables needed for plotting

```{R Define variables needed for plotting}
plot_theme = theme(
  plot.title = element_text(size = 26, face = "bold"),           
  plot.subtitle = element_text(size = 26),                       
  axis.title.x = element_text(size = 22),
  axis.title.y = element_text(size = 22),
  axis.text = element_text(size = 22),
  legend.text = element_text(size = 22),
  plot.background = element_rect(fill = 'white'),
  panel.background = element_rect(fill = 'white'),
  axis.line = element_line(color='black'),
  plot.margin = margin(2, 4, 2, 2, "cm"),
  panel.grid.major.x = element_blank(),
  panel.grid.major.y = element_line(colour = "grey", linewidth = 0.5),
)
```

### Correction factor plots

```{R}
plot_theme_CF = theme(
  plot.title = element_text(size = 26, face = "bold"),           
  plot.subtitle = element_text(size = 26),                       
  axis.title.x = element_text(size = 22),
  axis.title.y = element_text(size = 22),
  axis.text = element_text(size = 22),
  legend.text = element_text(size = 22),
  plot.background = element_rect(fill = 'white'),
  panel.background = element_rect(fill = 'white'),
  axis.line = element_line(color='black'),
  plot.margin = margin(2, 4, 2, 2, "cm"),
  panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_line(colour = "grey", linewidth = 0.5),
)
```

### Plot colors

```{R Define river colors}
river_colors <- c(
  "Meuse" = "#00be6c",
  "Lek canal" = "#BB9D00",
  "Overijsselse vecht" = "#00A5FF",
  "WWTP Werverschoof effluent canal" = "#7570b3",
  "WWTP Werverschoof\neffluent canal" = "#7570b3",
  "WWTP canal*"= "#7570b3",
  "Dommel" = "#F8766D",
  "Rhine" = "#e76bf3"
)
```

## Read in HC5 values

Read in PSSD++ from \@coffin2026 values for figures

```{R}
pssds <- read_excel(paste0(Input_dir, "/Coffin_PSSDs.xlsx")) |>
  mutate(particles_per_m3 = particles_per_L*1000) 

# Select only tier 3 values
tier3 <- pssds |>
  filter(Tier == 3) |>
  select(-c(Tier, ERM, Environment))

# Make a list of tier 3 values
tier3_list <- tier3 |>
  group_by(Method) |>
  summarise(metrics = list(
    setNames(particles_per_m3, Metric)
  )) |>
  deframe()
```

## Function for violin plot of concentrations compared to HC5

```{R}
options <- c("Rescaled (v_o2)", "Rescaled (v_o3)", "Reported concentrations") 
option_labels <- c("concentrationOption2" = "Uncertainty in alpha", "concentrationOption3" = "Uncertainty in alpha and detection limit", "reported" = "Reported concentrations") 

plot_facet_option <- function(
  df,
  option,
  tier_list,
  option_labels,
  plot_theme,
  tier_select = c("Volume ERM", "Surface area ERM")
) {
  # Data selecteren
  if (option %in% c("concentrationOption2", "concentrationOption3")) {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_violin(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        fill = River,
        color = River
      ),
      color = NA
    )
    xlabel = expression("Rescaled concentration (#/m"^3*")")
  } else if (option == "reported") {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_point(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        color = River
      ),
      shape = 16,
      size = 3
    )
    xlabel = expression("Reported concentration (#/m"^3*")")
  } else {
    stop("Onbekende optie")
  }

  facet_title <- option_labels[[option]]

  p <- ggplot2::ggplot(data = data_plot)

  # Gearceerde gebieden toevoegen
  if ("Volume ERM" %in% tier_select) {
    p <- p +
      ggplot2::annotate(
        "rect",
        xmin = tier_list$`Food dilution`["Q5"],
        xmax = tier_list$`Food dilution`["Q95"],
        ymin = -Inf,
        ymax = Inf,
        fill = ggplot2::alpha("blue", 0.1),
        color = NA
      )
  }
  if ("Surface area ERM" %in% tier_select) {
    p <- p +
      ggplot2::annotate(
        "rect",
        xmin = tier_list$`Tissue translocation`["Q5"],
        xmax = tier_list$`Tissue translocation`["Q95"],
        ymin = -Inf,
        ymax = Inf,
        fill = ggplot2::alpha("purple", 0.1),
        color = NA
      )
  }

  # Voeg de datalaag en rest toe
  p <- p +
    geom_layer +
    scale_fill_manual(values = river_colors) +
    scale_color_manual(values = river_colors) +
    scale_x_log10(
      breaks = scales::breaks_log(base = 10),
      labels = scales::label_log(base = 10)
    ) +
    ggplot2::labs(
      subtitle = paste0(tier_select, ", HC5"),
      x = xlabel,
      y = "Sample"
    ) +
    plot_theme +
    ggplot2::theme(
      legend.title = ggplot2::element_blank(),
      legend.position = "bottom",
      axis.text.x = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(size = 20), # grotere facet titel
      axis.ticks.y = ggplot2::element_blank(),
      panel.grid.minor.y = ggplot2::element_blank(),
      plot.margin = ggplot2::margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
    ) +
    ggplot2::coord_flip() +
    guides(
      color = guide_legend(override.aes = list(size = 4)),
      fill = guide_legend(override.aes = list(size = 4))
    )

  # Voeg vlines toe
  if ("Volume ERM" %in% tier_select) {
    p <- p +
      ggplot2::geom_vline(xintercept = tier_list$`Food dilution`["Median"], color = "blue", size = 0.6) +
      ggplot2::geom_vline(xintercept = tier_list$`Food dilution`["Q5"], linetype = "dashed", color = ggplot2::alpha("blue", 0.4)) +
      ggplot2::geom_vline(xintercept = tier_list$`Food dilution`["Q95"], linetype = "dashed", color = ggplot2::alpha("blue", 0.4))
  }
  if ("Surface area ERM" %in% tier_select) {
    p <- p +
      ggplot2::geom_vline(xintercept = tier_list$`Tissue translocation`["Median"], color = "purple", size = 0.6) +
      ggplot2::geom_vline(xintercept = tier_list$`Tissue translocation`["Q5"], linetype = "dashed", color = ggplot2::alpha("purple", 0.4)) +
      ggplot2::geom_vline(xintercept = tier_list$`Tissue translocation`["Q95"], linetype = "dashed", color = ggplot2::alpha("purple", 0.4))
  }

  if (option != "reported") {
      median_df <- df |>
        group_by(Sample_ID, Data_source) |>
        summarise(concentration = median(concentration))
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(median_df, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 8, # asterisk
      size = 2,
      stroke = 0.4,
      position = ggplot2::position_nudge(x = 0.6),
      show.legend = FALSE
    )
  }
    
  if (option == "reported") {
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(data_plot, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 8, # asterisk
      size = 2,
      stroke = 0.4,
      position = ggplot2::position_nudge(x = 0.2),
      show.legend = FALSE
    )
  }

  return(p)
}
```

## Function for violin plot of concentrations

```{R}
plot_facet_option_simple <- function(
  df,
  option,
  option_labels,
  plot_theme
) {
  # Data selecteren
  if (option %in% c("concentrationOption2", "concentrationOption3")) {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_violin(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        fill = River
      )
    )
  } else if (option == "reported") {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_point(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        color = River
      ),
      shape = 16,
      size = 3
    )
  } else {
    stop("Onbekende optie")
  }

  if(option == "reported"){
    scale_layer <- ggplot2::scale_color_manual(values = river_colors)
    xlabel = expression("Reported concentration (#/m"^3*")")
  } else {
    scale_layer <- ggplot2::scale_fill_manual(values = river_colors)
    xlabel = expression("Rescaled concentration (#/m"^3*")")
  }

  facet_title <- option_labels[[option]]

  p <- ggplot2::ggplot(data = data_plot) +
    geom_layer +
    scale_layer +
    scale_x_log10(
      breaks = scales::breaks_log(base = 10),
      labels = scales::label_log(base = 10)
    ) +
    ggplot2::labs(
      #title = facet_title,
      x = xlabel,
      y = "Sample"
    ) +
    plot_theme +
    ggplot2::theme(
      legend.title = ggplot2::element_blank(),
      legend.position = "bottom",
      axis.text.x = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(size = 13),
      axis.ticks.y = ggplot2::element_blank(),
      panel.grid.minor.y = ggplot2::element_blank(),
      plot.margin = ggplot2::margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
    ) +
    ggplot2::coord_flip()+
    guides(
      color = guide_legend(override.aes = list(size = 4)),
      fill = guide_legend(override.aes = list(size = 4))
    )
  
    if (option != "reported") {
      median_df <- df |>
        group_by(Sample_ID, Data_source) |>
        summarise(concentration = median(concentration))
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(median_df, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 8, # asterisk
      size = 2,
      stroke = 0.4,
      position = ggplot2::position_nudge(x = 0.6),
      show.legend = FALSE
    )
  }
    
  if (option == "reported") {
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(data_plot, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 8, # asterisk
      size = 2,
      stroke = 0.4,
      position = ggplot2::position_nudge(x = 0.2),
      show.legend = FALSE
    )
  }
  
  return(p)
}
```

# Figures

## Paragraph 3.2.1

```{R}
plot_data_without_effluent <- ConcAlphaAlignedUncertain_L |>
  filter(River != "WWTP Werverschoof effluent canal")

plot_data_effluent_only <- ConcAlphaAlignedUncertain_L |>
  filter(River == "WWTP Werverschoof effluent canal")
```

### Reported concentrations without effluent

```{R}
option = "reported"
p = plot_facet_option_simple(
  df = plot_data_without_effluent,
  option = option,
  option_labels = option_labels,
  plot_theme = theme_minimal())

print(p)

#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_Rescaled concentration violin ", option, " Background", ".png"), plot = p, width = 12, height = 9, dpi = 300)
```

### Reported concentrations effluent canal only

```{R}
p = plot_facet_option_simple(  
  df = plot_data_effluent_only,
  option = "reported",
  option_labels = option_labels,
  plot_theme = theme_minimal())

print(p)

#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_Rescaled concentration violin ", option, " effluent canal", ".png"), plot = p, width = 12, height = 9, dpi = 300)
```

## Paragraph 3.2.2

### Power law slope

```{R}
ConcAlphaUncertain_plot <- 
  fAlignUncertain(
    ConcAlphaCombined = ConcPSD, 
    nRuns = 1000,
    SizeParameter = "longest") 

ConcAlphaUncertain_plot$River <- gsub(
  "WWTP Werverschoof effluent canal",
  "WWTP Werverschoof\neffluent canal",
  ConcAlphaUncertain_plot$River
) 

ConcAlphaUncertain_plot$Analysis_method <- gsub(
  "ATR-FTIR and Micro-FTIR",
  "FTIR",
  ConcAlphaUncertain_plot$Analysis_method
) 

triangle_alpha_plot_river <- ggplot(ConcAlphaUncertain_plot, 
                             aes(x = triangle_alpha, 
                                 y = forcats::fct_reorder(River, triangle_alpha, .fun = median), 
                                 fill = River)) +
  geom_violin() +
  facet_wrap(~ Analysis_method) + 
  theme_minimal() +
  scale_fill_manual(values = river_colors) +
  theme(legend.position = "none",
            strip.text = element_text(size = 22)) +
  labs(
    x="Power law slope",
    y=""
  ) 

asterisk_data <- ConcAlphaUncertain_plot %>%
  filter(River %in% c("Lek canal", "Rhine")) %>%
  group_by(River, Analysis_method) %>%
  summarise(triangle_alpha = min(triangle_alpha), .groups = 'drop')

triangle_alpha_plot_river <- triangle_alpha_plot_river +
  ggplot2::geom_point(
    data = asterisk_data,
    mapping = ggplot2::aes(
      x = triangle_alpha,
      y = forcats::fct_reorder(River, triangle_alpha, .fun = median)
    ),
    color = "black",
    shape = 8, # asterisk
    size = 3,
    stroke = 0.6,
    position = ggplot2::position_nudge(x = -0.2),
    show.legend = FALSE
  )

triangle_alpha_plot_river

#ggsave(paste0(project_dir, alphas_dir, "/", Sys.Date(), "_Triange_alpha_river_analysis_method", ".png"), plot = triangle_alpha_plot_river, width = 12, height = 6, dpi = 300)
```

### Correction factor

```{R}
CF_plotdata <- ConcAlphaAlignedUncertain |>
  mutate(
    Analysis_method = case_when(
      Analysis_method == "ATR-FTIR and Micro-FTIR" ~ "FTIR",
      TRUE ~ Analysis_method
    ),
    CF_factor = paste0(River, ", ", Analysis_method),
    CF_factor = fct_reorder(CF_factor, CF_option3, .fun = median, .desc = FALSE)
  )

cf_plot <- ggplot(CF_plotdata, aes(x = CF_option3, y = CF_factor, fill = River)) +
  geom_violin() +
  scale_fill_manual(values = river_colors) +
  scale_x_log10(labels = scales::comma) +
  theme_minimal() +
  theme(
    legend.position = "none"
  ) +
  labs(
    x = "Correction factor",
    y = ""
  )

asterisk_data <- CF_plotdata %>%
  filter(River %in% c("Lek canal", "Rhine")) %>%
  group_by(CF_factor, River) %>%  
  summarise(CF_option3 = min(CF_option3), .groups = 'drop')

cf_plot <- cf_plot +
  ggplot2::geom_point(
    data = asterisk_data,
    mapping = ggplot2::aes(
      x = CF_option3,
      y = forcats::fct_reorder(CF_factor, CF_option3, .fun = median)
    ),
    color = "black",
    shape = 8, # asterisk
    size = 3,
    stroke = 0.6,
    position = ggplot2::position_nudge(x = -0.2),
    show.legend = FALSE
  )

cf_plot

#ggsave(paste0(project_dir, alphas_dir, "/", Sys.Date(), "_Correction_factors", ".png"),width = 14, height = 10, dpi = 300)
```

## Paragraph 3.2.3

### Spread in average concentrations

```{R}
all_rescaled_concentrations <- rbind(plot_data_effluent_only, plot_data_without_effluent) |> 
  filter(OptionType == "concentrationOption3")

spread_river <- ggplot(all_rescaled_concentrations, 
                             aes(x = concentration, 
                                 y = forcats::fct_reorder(River, concentration, .fun = median), 
                                 fill = River)) +
  geom_violin() +
  facet_wrap(~ Analysis_method) + 
  scale_fill_manual(values = river_colors) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    x = expression("Rescaled particle number concentration (#/m"^3*")"),
    y = ""
  ) +
  scale_x_log10(
    breaks = scales::breaks_log(base = 10),
    labels = scales::label_log(base = 10)
  )

asterisk_data <- all_rescaled_concentrations %>%
  filter(River %in% c("Lek canal", "Rhine")) %>%
  group_by(River, Analysis_method) %>%  
  summarise(concentration = min(concentration), .groups = 'drop')

spread_river <- spread_river +
  ggplot2::geom_point(
    data = asterisk_data,
    mapping = ggplot2::aes(
      x = concentration,
      y = forcats::fct_reorder(River, concentration, .fun = median)
    ),
    color = "black",
    shape = 8, # asterisk
    size = 3,
    stroke = 0.6,
    position = ggplot2::position_nudge(x = -1),
    show.legend = FALSE
  )

spread_river

#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_Spread comparison concentration violin ", ".png"), plot = spread_river, width = 16, height = 9, dpi = 300)
```

## Paragraph 3.5

### Without effluent

```{R, warning=F, message=F}
options <- c("concentrationOption3")

for(option in options){
  p <- plot_facet_option(
  df = plot_data_without_effluent,
  option = option,
  tier_list = tier3_list,
  tier_select = "Volume ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, "_Food Dilution background.png"), plot = p, width = 12, height = 9, dpi = 300)
}

for(option in options){
  p <- plot_facet_option(
  df = plot_data_without_effluent,
  option = option,
  tier_list = tier3_list,
  tier_select = "Surface area ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir,concentrations_dir, "/", Sys.Date(), "_RescaledConcentrationHC5Tier3_", option, " Tissue Translocation background", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}
```

### Effluent only

```{R, warning=F, message=F}
options <- c("concentrationOption3")

option = "concentrationOption3"
for(option in options){
  p <- plot_facet_option(
  df = plot_data_effluent_only,
  option = option,
  tier_list = tier3_list,
  tier_select = "Volume ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, " Food dilution effluent canal", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}

for(option in options){
  p <- plot_facet_option(
  df = plot_data_effluent_only,
  option = option,
  tier_list = tier3_list,
  tier_select = "Surface area ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, " Tissue translocation effluent canal", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}
```

## Paragraph 3.2.4

### Polymer fraction comparison

#### Load modelled concentrations

```{R Load data, warning=F, message=F}
Concentrations_2019 <- read_rds(paste0(project_dir, output_dir, "Data/Particle_concentrations_SimpleBox.rds"))

# sum air and cloudwater
Concentrations_2019 <- Concentrations_2019 |>
  filter(Scale == "Regional") |>
  filter(SubCompart == "river") |>
  filter(Polymer != "General")

unique(Concentrations_2019$Polymer)

concentrations_SB_polymer_mapped <- Concentrations_2019 |>
  mutate(Polymer = case_when(
    Polymer == "OTHER" ~ "Others",
    Polymer %in% c("LDPE", "HDPE") ~ "PE",
    Polymer == "RUBBER" ~ "Other synthetic rubbers",
    TRUE ~ Polymer
  )) |>
  group_by(RUN, Unit, Concentration, Polymer, Year, Scale, SubCompart) |>
  summarise(Concentration = sum(Concentration))

Mean_SB_concentrations <- concentrations_SB_polymer_mapped |>
    group_by(Unit, Polymer, Year, Scale, SubCompart) |>
    summarise(concentration = mean(Concentration)) |>
    ungroup()

SB_total_concentration <- sum(Mean_SB_concentrations$concentration)

SB_polymer_fractions <- Mean_SB_concentrations %>%
  mutate(mean_fraction = concentration/SB_total_concentration) |>
  mutate(mean_percentage = mean_fraction*100) |>
  mutate(Data_source = "SimpleBox*") |>
  select(Data_source, Polymer, mean_fraction, mean_percentage)
```

#### Load particle measurement concentrations

```{R, warning=F, message=F}
ReportedPolymerData <- readRDS(paste0(project_dir, output_dir, "Data/", DataVersionConc, "Particle_concentrations_per_polymer.rds")) 

particle_concentrations_polymer <- ReportedPolymerData |>
  group_by(Data_source, Polymer) |>
  summarise(concentration = mean(concentration)) |>
  ungroup() |>
  mutate(Data_source = case_when(
    Data_source == "Mintenig, Hunnestad & Koelmans (2025)" ~ "Mintenig et al. (2025)",
    TRUE ~ Data_source
  ))

particle_concentrations_total_calculated <- particle_concentrations_polymer |>
  group_by(Data_source) |>
  summarise(total_concentration = sum(concentration)) |>
  ungroup() 

particle_concentrations_polymer <- particle_concentrations_polymer |>
  left_join(particle_concentrations_total_calculated) |>
  group_by(Data_source, Polymer) |>
  summarise(
    mean_fraction = mean(concentration / total_concentration),
    .groups = "drop"
  ) |>
  mutate(mean_percentage = mean_fraction*100) 

plot_data <- rbind(particle_concentrations_polymer, SB_polymer_fractions)

unique(particle_concentrations_polymer$Polymer)

controle <- plot_data %>%
  group_by(Data_source) %>%
  summarise(som_mean_fraction = sum(mean_fraction))

print(controle)

PolymerMeasurements <-
  ReportedPolymerData |> ungroup() |> 
  group_by(River, Polymer, Analysis_method) |> 
  summarise(PolConcAvg = mean(concentration)) |>
  left_join(
    ReportedPolymerData |> ungroup() |> 
      group_by(River, Polymer, Analysis_method) |> 
      summarise(PolConcAvg = mean(concentration)) |>
      ungroup() |> 
      group_by(River,Analysis_method) |>
      summarise(TotConcAvg = sum(PolConcAvg)) |>
      ungroup() 
  ) |> mutate(PolFrac = PolConcAvg/TotConcAvg)

PolymerIndividualMeasurements <- ReportedPolymerData |> 
  ungroup() |> 
  group_by(River, Polymer, Analysis_method,Sample_ID) |> 
  summarise(PolConcAvg = mean(concentration)) |>
  left_join(
    ReportedPolymerData |> 
      ungroup() |> 
      group_by(River, Polymer, Analysis_method,Sample_ID) |> 
      summarise(PolConcAvg = mean(concentration)) |>
      ungroup() |> 
      group_by(River,Analysis_method) |>
      summarise(TotConcAvg = sum(PolConcAvg)) |>
      ungroup()) |> 
  mutate(PolFrac = PolConcAvg/TotConcAvg)
```

#### Plot

```{R, warning=F, message=F}
n_polymeren <- length(unique(plot_data$Polymer))

p1 <- ggplot(plot_data, aes(x = Data_source, y = mean_percentage, fill = Polymer)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_viridis_d(option = "turbo", end = 0.9) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "",
    y = "",
    fill = ""
  ) +
  theme_minimal() + 
  coord_flip() +
  theme_minimal()

print(p1)

#ggsave(paste0(project_dir,"Figures/Measurements/Polymer_distributions_compared.png"), plot = p1, width = 12, height = 9, dpi = 1200)
```

#### Save table with average fractions

```{R, warning=F, message=F}
polymer_fraction_table <- plot_data |>
  select(-mean_percentage) |>
  pivot_wider(
    id_cols = Polymer,
    names_from = Data_source,
    values_from = mean_fraction)

# openxlsx::write.xlsx(polymer_fraction_table, file = paste0(project_dir, "Tables/Polymer_share_per_source.xlsx"))
```

## Appendix 4

### Reported concentrations compared to HC5

#### Without effluent

```{R, warning=F, message=F}
options <- c("reported")

for(option in options){
  p <- plot_facet_option(
  df = plot_data_without_effluent,
  option = option,
  tier_list = tier3_list,
  tier_select = "Volume ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, "_Food Dilution background.png"), plot = p, width = 12, height = 9, dpi = 300)
}

for(option in options){
  p <- plot_facet_option(
  df = plot_data_without_effluent,
  option = option,
  tier_list = tier3_list,
  tier_select = "Surface area ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir,concentrations_dir, "/", Sys.Date(), "_RescaledConcentrationHC5Tier3_", option, " Tissue Translocation background", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}
```

#### Effluent only

```{R, warning=F, message=F}
options <- c("reported")

option = "concentrationOption3"
for(option in options){
  p <- plot_facet_option(
  df = plot_data_effluent_only,
  option = option,
  tier_list = tier3_list,
  tier_select = "Volume ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, " Food dilution effluent canal", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}

for(option in options){
  p <- plot_facet_option(
  df = plot_data_effluent_only,
  option = option,
  tier_list = tier3_list,
  tier_select = "Surface area ERM",
  option_labels = option_labels,
  plot_theme = theme_minimal()
)
  print(p)
  
#ggsave(paste0(project_dir, concentrations_dir, "/",Sys.Date(),"_RescaledConcentrationHC5Tier3_", option, " Tissue translocation effluent canal", ".png"), plot = p, width = 12, height = 9, dpi = 300)
}
```

### Polypropylene figures

#### Read in data

```{R, warning=F, message=F}
AlphasPPLDIR <- read_csv(paste0(project_dir, output_dir,"Data/powerlawFitLDIR_pp", DataVersionPowerlawLDIR, ".csv")) |>
  select(-`...1`)
AlphasPPFTIR <-  read_csv(paste0(project_dir, output_dir,"Data/powerlawFitFTIR_pp", DataVersionPowerlawFTIR, ".csv")) |>
  select(-`...1`)

concentrations_PP <- read_rds(paste0(project_dir, output_dir,"Data/",DataVersionConc,"Particle_concentrations_per_polymer.rds")) |>
  filter(Polymer == "PP")

# Exclude samples from Mintenig (2020) which were indicated only for MP > 300 um.
concentrations_PP <- concentrations_PP |>
  filter(!endsWith(Location_name,"*"))

alphas <- AlphasPPFTIR|>
  mutate(Analysis_method = "ATR-FTIR and Micro-FTIR")|>
  bind_rows(AlphasPPLDIR|>
              mutate(Analysis_method = "LDIR"))
```

#### Align concentrations

```{R, warning=F, message=F}
ConcPSD_PP <- concentrations_PP |>
  filter(!is.na(River))|>
  left_join(alphas |> # leaves out the ones where id == All
              select(id,
                     param,
                     alpha,
                     alpha_lower,
                     alpha_upper,
                     Analysis_method),
            by = c("River" = "id",
                   "Analysis_method" = "Analysis_method"),
            relationship = "many-to-many")|>
  filter(!Analysis_method=="Light microscopy")

ConcAlphaUncertain_PP <- 
  fAlignUncertain(
    ConcAlphaCombined = ConcPSD_PP, 
    nRuns = 1000,
    SizeParameter = "longest")

ConcAlphaAlignedUncertain_PP <- ConcAlphaUncertain_PP |> 
  mutate(CF_option2 = 
           F.Correctionfactor(Upper_sampling_size_limit_um,
                              Lower_sampling_size_limit_um,
                              5000,
                              1,
                              triangle_alpha))|>
  mutate(concentrationOption2 = CF_option2*
           concentration) |> 
  mutate(CF_option3 = 
           F.Correctionfactor(triangle_sampling_upper,
                              triangle_sampling_lower,
                              5000,
                              1,
                              triangle_alpha))|>
  mutate(concentrationOption3 = CF_option3*
           concentration)

ConcAlphaAlignedUncertain_PP_L <-
  ConcAlphaAlignedUncertain_PP |> 
  rename(reported = concentration) |> 
  pivot_longer(
    cols = c(reported,concentrationOption2,concentrationOption3),
    names_to = "OptionType",
    values_to = "concentration"
  )

rm(ConcAlphaUncertain_PP)
```

```{R Save rescaled PP concentrations, eval=FALSE, include=FALSE}
saveRDS(ConcAlphaAlignedUncertain_PP_L, paste0(project_dir, output_dir, "Data/Rescaled_concentrations_PP_", DataVersionPowerlaw, ".Rds"))
```

#### Plot

```{R Plot, eval=FALSE, warning=F, message=F}
plot_facet_option_simple <- function(
  df,
  option,
  option_labels,
  plot_theme
) {
  # Data selecteren
  if (option %in% c("concentrationOption2", "concentrationOption3")) {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_violin(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        fill = River
      )
    )
  } else if (option == "reported") {
    data_plot <- dplyr::filter(df, OptionType == option)
    geom_layer <- ggplot2::geom_point(
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median),
        color = River
      ),
      shape = 16,
      size = 3
    )
  } else {
    stop("Onbekende optie")
  }

  if(option == "reported"){
    scale_layer <- ggplot2::scale_color_manual(values = river_colors)
    xlabel = expression("Reported concentration (#/m"^3*")")
  } else {
    scale_layer <- ggplot2::scale_fill_manual(values = river_colors)
    xlabel = expression("Rescaled concentration (#/m"^3*")")
  }

  facet_title <- option_labels[[option]]

  p <- ggplot2::ggplot(data = data_plot) +
    geom_layer +
    scale_layer +
    scale_x_log10(
      breaks = scales::breaks_log(base = 10),
      labels = scales::label_log(base = 10)
    ) +
    ggplot2::labs(
      #title = facet_title,
      x = xlabel,
      y = "Sample"
    ) +
    plot_theme +
    ggplot2::theme(
      legend.title = ggplot2::element_blank(),
      legend.position = "bottom",
      axis.text.x = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(size = 13),
      axis.ticks.y = ggplot2::element_blank(),
      panel.grid.minor.y = ggplot2::element_blank(),
      plot.margin = ggplot2::margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
    ) +
    ggplot2::coord_flip()+
    guides(
      color = guide_legend(override.aes = list(size = 4)),
      fill = guide_legend(override.aes = list(size = 4))
    )
  
    if (option != "reported") {
      median_df <- df |>
        group_by(Sample_ID, Data_source) |>
        summarise(concentration = median(concentration))
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(median_df, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 3,
      size = 3,
      stroke = 0.8,
      position = ggplot2::position_nudge(x = 0.3),
      show.legend = FALSE
    )
  }
    
  if (option == "reported") {
    p <- p + ggplot2::geom_point(
      data = dplyr::filter(data_plot, Data_source == "Mughini-Gras et al. (2021)"),
      mapping = ggplot2::aes(
        x = concentration,
        y = forcats::fct_reorder(Sample_ID, concentration, .fun = median)
      ),
      color = "black",
      shape = 3,
      size = 3,
      stroke = 0.8,
      position = ggplot2::position_nudge(x = 0.3),
      show.legend = FALSE
    )
  }
  
  return(p)
}
```

```{R Plot PP concentrations, warning=F, message=F}
plot_data_meuse_rhine <- ConcAlphaAlignedUncertain_PP_L |>
  filter(River %in% c("Meuse", "Rhine"))

plot_data_other <- ConcAlphaAlignedUncertain_PP_L |>
  filter(!River %in% c("Meuse", "Rhine")) |>
  mutate(River = case_when(
    River == "WWTP Werverschoof effluent canal" ~ "WWTP canal*",
    TRUE ~ River
  ))

option = "concentrationOption3"

# Meuse and Rhine
p = plot_facet_option_simple(
  df = plot_data_meuse_rhine,
  option = option,
  option_labels = option_labels,
  plot_theme = theme_minimal())

print(p)

#ggsave(paste0(project_dir, concentrations_dir, "/", Sys.Date(), "_Rescaled concentration violin PP Meuse Rhine ", option, " Background", ".png"), plot = p, width = 12, height = 9, dpi = 300)

# Other water bodies
p = plot_facet_option_simple(
  df = plot_data_other,
  option = option,
  option_labels = option_labels,
  plot_theme = theme_minimal())

print(p)

#ggsave(paste0(project_dir, concentrations_dir, "/", Sys.Date(), "_Rescaled concentration violin PP Other ", option, " Background", Sys.Date(), ".png"), plot = p, width = 12, height = 9, dpi = 300)
```

#### Table with summary statistics

```{R, warning=F, message=F}
PP_averages_table <- ConcAlphaAlignedUncertain_PP_L |>
  filter(OptionType == "concentrationOption3") |>
  group_by(Sample_ID, River) |>
  summarise(concentration = mean(concentration)) |>
  ungroup() |>
  group_by(River) |>
summarise(
  min = format(min(concentration), scientific = TRUE, digits = 2),
  max = format(max(concentration), scientific = TRUE, digits = 2),
  mean = format(mean(concentration), scientific = TRUE, digits = 2)
) |>
  ungroup() |>
  mutate(min = as.character(min),
         mean = as.character(mean), 
         max = as.character(max),
         `Concentration (#/m3)` = paste0(mean, " (", min, "-", max, ")")) |>
  select(River, `Concentration (#/m3)`) |>
  rename(`Measured (#/m3)` = `Concentration (#/m3)`)

#openxlsx::write.xlsx(PP_averages_table, file = paste0(output_dir, "Tables/Number_concentrations_PP.xlsx"))
```
